---
title: "10_Function_Factories"
output: pdf_document
date: "2023-01-31"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 10 Function factories
## 10.1 Introduction
```{r}
library(tidyverse)
```

```{r}
power1 <- function(exp) {
  function(x) {
    x ^ exp
  }
}

square <- power1(2)
cube <- power1(3)
```

```{r}
library(rlang)
library(ggplot2)
library(scales)
```

## 10.2 Factory fundamentals

```{r}
square

cube
```

```{r}
env_print(square)

env_print(cube)
```

```{r}
fn_env(square)$exp

fn_env(cube)$exp
```

### 10.2.2 Diagram conventions

```{r}
square(10)
```

### 10.2.3 Forcing evaluation
```{r}
x <- 2
square <- power1(x)
x <- 3
```

```{r}
square(2)
```

```{r}
power2 <- function(exp) {
  force(exp)
  function(x) {
    x ^ exp
  }
}

x <- 2
square <- power2(x)
x <- 3
square(2)
```

### 10.2.4 Stateful functions

```{r}
new_counter <- function() {
  i <- 0
  
  function() {
    i <<- i + 1
    i
  }
}

counter_one <- new_counter()
counter_two <- new_counter()
```

```{r}
counter_one()

counter_one()

counter_two()

counter_one()

counter_two()
```

### 10.2.5 Garbage collection

```{r}
f1 <- function(n) {
  x <- runif(n)
  m <- mean(x)
  function() m
}

g1 <- f1(1e6)
lobstr::obj_size(g1)
#> 8,013,104 B

f2 <- function(n) {
  x <- runif(n)
  m <- mean(x)
  rm(x)
  function() m
}

g2 <- f2(1e6)
lobstr::obj_size(g2)
```

#### 10.2.6 Exercises

1. The definition of `force()` is simple:

```{r}
force
```

Why is it better to force(x) instead of just x?

Using `force(x)` is better than `x` because it makes the evaluation of x occur immediately when the function is created rather than lazily executing the first time the function is called. If the value of `x` changes before the function is called, the value of `x` when the function was created will not be used and the new value of `x` will be used

```{r}
timeser <- function(multiplier){
  function(x) x * multiplier
}
x <- 2
doubler <- timeser(x)
# Function called before x changed
doubler(2)
x <- 3
doubler(2)

x <- 2
doubler <- timeser(x)
# Function not called before x changed
x <- 3
doubler(2)

# Forced operation
timeser2 <- function(multiplier){
  force(multiplier)
  function(x) x * multiplier
}
x <- 2
doubler <- timeser2(x)
# Function not called before x changed
x <- 3
doubler(2)
```

2. Base R contains two function factories, `approxfun()` and `ecdf()`. Read their documentation and experiment to figure out what the functions do and what they return.

```{r}
x <- 1:10
y <- rnorm(10)
par(mfrow = c(2,1))
plot(x, y, main = "approx(.) and approxfun(.)")
points(approx(x, y), col = 2, pch = "*")
points(approx(x, y, method = "constant"), col = 4, pch = "*")

f <- approxfun(x, y)
curve(f(x), 0, 11, col = "green2")
points(x, y)
is.function(fc <- approxfun(x, y, method = "const")) # TRUE
curve(fc(x), 0, 10, col = "darkblue", add = TRUE)
## different extrapolation on left and right side :
plot(approxfun(x, y, rule = 2:1), 0, 11,
     col = "tomato", add = TRUE, lty = 3, lwd = 2)
```

```{r}
x <- rnorm(12)
Fn <- ecdf(x)
Fn     # a *function*
Fn(x)  # returns the percentiles for x
tt <- seq(-2, 2, by = 0.1)
12 * Fn(tt) # Fn is a 'simple' function {with values k/12}
summary(Fn)
##--> see below for graphics
knots(Fn)  # the unique data values {12 of them if there were no ties}

y <- round(rnorm(12), 1); y[3] <- y[1]
Fn12 <- ecdf(y)
Fn12
knots(Fn12) # unique values (always less than 12!)
summary(Fn12)
summary.stepfun(Fn12)
```

3. Create a function `pick()` that takes an index, `i`, as an argument and returns a function with an argument `x` that subsets `x` with `i`.

```{r}
pick <- function(index){
  function(x) x[index]
}

a <- 1:10
b <- LETTERS[1:10]
pick_2s <- pick(c(2,4,8))
pick_2s(a)
pick_2s(b)
```

4. Create a function that creates functions that compute the $i^th$ central moment of a numeric vector. You can test it by running the following code:

```{r}
moment <- function(index){
  function(x){
    sum((x - mean(x)) ^ index) / length(x)
  }
}
m1 <- moment(1)
m2 <- moment(2)

x <- runif(100)
stopifnot(all.equal(m1(x), 0))
stopifnot(all.equal(m2(x), var(x) * 99 / 100))
```

5. What happens if you don’t use a closure? Make predictions, then verify with the code below.
```{r}
i <- 0
new_counter2 <- function() {
  i <<- i + 1
  i
}

new_counter2()
new_counter2()
i
i <- 9
new_counter2()
i
```
Uses the parent environment and changes when global `i` changes in this case
