---
title: "19 Quasiquotation"
output: html_document
date: "2023-07-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 19 Quasiquotation
## 19.1 Introduction

```{r}
library(rlang)
library(purrr)
```

## 19.2 Motivation
```{r}
paste("Good", "morning", "Hadley")

paste("Good", "afternoon", "Alice")
```

```{r}
cement <- function(...) {
  args <- ensyms(...)
  paste(purrr::map(args, as_string), collapse = " ")
}

cement(Good, morning, Hadley)

cement(Good, afternoon, Alice)
```

```{r}
name <- "Hadley"
time <- "morning"

paste("Good", time, name)
```

```{r}
cement(Good, time, name)
```

```{r}
cement(Good, !!time, !!name)
```

```{r}
paste("Good", time, name)
cement(Good, !!time, !!name)
```

### 19.2.1 Vocabulary

```{r, eval = FALSE}
library(MASS)

# fails
MASS
```
### 19.2.2 Exercises

1.For each function in the following base R code, identify which arguments are quoted and which are evaluated.

```{r, error = TRUE}
library(MASS) # MASS is quoted
MASS

mtcars2 <- subset(mtcars, cyl == 4) # mtcars is evaluated, and cyl is quoted
mtcars
cyl == 4

with(mtcars2, sum(vs)) # mtcars is evaluated and sum(vs) is quoted
mtcars2
sum(vs)

sum(mtcars2$am) #mtcars2$am is evaulated and am is quoted
mtcars2$am
$am

rm(mtcars2) # mtcars2 is quoted
```

2. For each function in the following tidyverse code, identify which arguments are quoted and which are evaluated.

```{r}
library(dplyr) # dplyr is quoted
library(ggplot2) #ggplot2 is quoted

by_cyl <- mtcars %>% # mtcars is evaluated
  group_by(cyl) %>% #cyl is quoted
  summarise(mean = mean(mpg)) # mean = mean(mpg) is quoted

ggplot(by_cyl, aes(cyl, mean)) + geom_point() # by_cyl evaluated, aes is evaluated, cyl and mean are quoted
```

## 19.3 Quoting
### 19.3.1 Capturing expressions

```{r}
expr(x + y)

expr(1 / 2 / 3)
```

```{r}
f1 <- function(x) expr(x)
f1(a + b + c)
```

```{r}
f2 <- function(x) enexpr(x)
f2(a + b + c)
```

```{r}
f <- function(...) enexprs(...)
f(x = 1, y = 10 * z)
```

```{r}
exprs(x = x ^ 2, y = y ^ 3, z = z ^ 4)
```
### 19.3.2 Capturing symbols
```{r}
f <- function(...) ensyms(...)
f(x)


f("x")
```

### 19.3.3 With base R

```{r}
quote(x + y)
```

```{r}
f3 <- function(x) substitute(x)
f3(x + y)
```

```{r}
alist(x = 1, y = x + 2)
```

```{r}
f <- function(...) as.list(substitute(...()))
f(x = 1, y = 10 * z)
```

### 19.3.4 Substitution
```{r}
f4 <- function(x) substitute(x * 2)
f4(a + b + c)
```

```{r}
substitute(x * y * z, list(x = 10, y = quote(a + b)))
```

### 19.3.5 Summary
### 19.3.6 Exercises

1. How is `expr()` implemented? Look at its source code.
```{r}
expr
```
It's a wrapper for `enexpr`

2. Compare and contrast the following two functions. Can you predict the output before running them?

```{r}
f1 <- function(x, y) {
  exprs(x = x, y = y)
} # captures x = x and y = y
f2 <- function(x, y) {
  enexprs(x = x, y = y)
} # captures the user input en is for users input therefore we'd get x = a + b and y = c + d
f1(a + b, c + d)
f2(a + b, c + d)
```

3. What happens if you try to use `enexpr()` with an expression (i.e.  `enexpr(x + y)` ? What happens if `enexpr()` is passed a missing argument?

```{r, error = T}
enexpr(x + y)
```
You get an error, the argument has to be a symbol

```{r, error = T}
enexpr(x = )
```
You get an error for an unused argument

4. How are `exprs(a)` and `exprs(a = )` different? Think about both the input and the output.

```{r}
exprs(a)

exprs(a = )
```
First one you get a list with the value "a" in the second you get a list with an object "a" which doesn't have a value

5. What are other differences between `exprs()` and `alist()`? Read the documentation for the named arguments of `exprs()` to find out.

.named 	If TRUE, unnamed inputs are automatically named with as_label()  

.ignore_empty Whether to ignore empty arguments. Can be one of "trailing", "none", "all". If "trailing"

6. The documentation for `substitute()` says:

Substitution takes place by examining each component of the parse tree as follows:

* If it is not a bound symbol in env, it is unchanged.
* If it is a promise object (i.e., a formal argument to a function) the expression slot of the promise replaces the symbol.
* If it is an ordinary variable, its value is substituted, unless env is .GlobalEnv in which case the symbol is left unchanged.

```{r}
substitute(x + y, list(x = 2)) # y does not change

substitute(mean(x, ...), list(x = 1, na.rm = T)) # spits back the function with the symbols replaced

y <- 5
substitute(x + y, .GlobalEnv) # doesn't change y
```

## 19.4 Unquoting
### 19.4.1 Unquoting one argument

```{r}
x <- expr(-1)
expr(f(!!x, y))
```

```{r}
a <- sym("y")
b <- 1
expr(f(!!a, !!b))
```

```{r}
mean_rm <- function(var) {
  var <- ensym(var)
  expr(mean(!!var, na.rm = TRUE))
}
expr(!!mean_rm(x) + !!mean_rm(y))
```

```{r}
x1 <- expr(x + 1)
x2 <- expr(x + 2)

expr(!!x1 / !!x2)
```

### 19.4.2 Unquoting a function

```{r}
f <- expr(foo)
expr((!!f)(x, y))
```

```{r}
f <- expr(pkg::foo)
expr((!!f)(x, y))
```

```{r}
f <- expr(pkg::foo)
call2(f, expr(x), expr(y))
```

### 19.4.3 Unquoting a missing argument
```{r, error = TRUE}
arg <- missing_arg()
expr(foo(!!arg, !!arg))
```

```{r}
expr(foo(!!maybe_missing(arg), !!maybe_missing(arg)))
```

### 19.4.4 Unquoting in special forms
```{r, error = TRUE}
expr(df$!!x)
```

```{r}
x <- expr(x)
expr(`$`(df, !!x))
```

### 19.4.5 Unquoting many arguments

```{r}
xs <- exprs(1, a, -b)
expr(f(!!!xs, y))
```

```{r}
ys <- set_names(xs, c("a", "b", "c"))
expr(f(!!!ys, d = 4))
```

```{r}
call2("f", !!!xs, expr(y))
```

### 19.4.6 The polite fiction of !!

```{r}
!!TRUE
!!!TRUE
```

```{r, error = TRUE}
x <- quote(variable)
!!x
```

```{r}
# BAD
df <- data.frame(x = 1:5)
y <- 100
with(df, x + !!y)
```

### 19.4.7 Non-standard ASTs
```{r}
x1 <- expr(class(!!data.frame(x = 10)))
x1
eval(x1)
```

```{r}
expr_print(x1)

lobstr::ast(!!x1)
```

```{r}
x2 <- expr(f(!!c(1L, 2L, 3L, 4L, 5L)))
x2

expr_print(x2)

lobstr::ast(!!x2)
```

```{r}
x3 <- expr(1 + !!expr(2 + 3))
x3

lobstr::ast(!!x3)
```

### 19.4.8 Exercises
1. Given the following components:, Use quasiquotation to construct the following calls:
```{r}
xy <- expr(x + y)
xz <- expr(x + z)
yz <- expr(y + z)
abc <- exprs(a, b, c)
```

```{r}
# (x + y) / (y + z)
expr(!!xy / !!yz)

#-(x + z) ^ (y + z)
expr(-`^`(!!xz, !!yz))

# (x + y) + (y + z) - (x + y)
expr(((!!xy)) + !!yz - !!xy)

#atan2(x + y, y + z)
expr(atan2(!!xy, !!yz))

#sum(x + y, x + y, y + z)
expr(sum(!!xy, !!xy, !!yz))

#sum(a, b, c)
expr(sum(!!!abc))

#mean(c(a, b, c), na.rm = TRUE)
expr(mean(c(!!!abc), na.rm = TRUE))

#foo(a = x + y, b = y + z)
expr(foo(a = !!xy, b = !!yz))
```

2. The following two calls print the same, but are actually different:

```{r}
(a <- expr(mean(1:10)))

(b <- expr(mean(!!(1:10))))

(c <- expr(mean(!!1:10)))

identical(a, b)
identical(a,c)
mean(1:10)
mean(!!(1:10))
```

The first is taking the mean of 1:10, while the second is taking the mean of a vector TRUEs. The second takes 1:10, negates it, were TRUE, now they're FALSE, then negates that to all TRUE. The presence of the `()` causes this

## 19.5 Non-quoting

```{r}
xyz <- bquote((x + y + z))
bquote(-.(xyz) / 2)
```

```{r}
x <- list(var = 1, y = 2)
var <- "y"

x$var

x[[var]]
```

```{r}
x <- 1
rm(x)

y <- 2
vars <- c("y", "vars")
rm(list = vars)
```

```{r}
# Shows help for var
help(var)

var <- "mean"
# Shows help for mean
help(var)

var <- 10
# Shows help for var
help(var)
```

```{r}
palette(RColorBrewer::brewer.pal(3, "Set1"))
plot(
  Sepal.Length ~ Petal.Length, 
  data = iris, 
  col = Species, 
  pch = 20, 
  cex = 2
)
```

## 19.6 ... (dot-dot-dot)

```{r}
dfs <- list(
  a = data.frame(x = 1, y = 2),
  b = data.frame(x = 3, y = 4)
)
```

```{r}
var <- "x"
val <- c(4, 3, 9)
```

```{r}
dplyr::bind_rows(!!!dfs)
```

```{r}
tibble::tibble(!!var := val)
```

```{r, error = TRUE}
tibble::tibble(!!var = value)
```

### 19.6.1 Examples

```{r}
set_attr <- function(.x, ...) {
  attr <- rlang::list2(...)
  attributes(.x) <- attr
  .x
}

attrs <- list(x = 1, y = 2)
attr_name <- "z"

1:10 %>%
  set_attr(w = 0, !!!attrs, !!attr_name := 3) %>% 
  str()
```

### 19.6.2 exec()

```{r}
# Directly
exec("mean", x = 1:10, na.rm = TRUE, trim = 0.1)

# Indirectly
args <- list(x = 1:10, na.rm = TRUE, trim = 0.1)
exec("mean", !!!args)

# Mixed
params <- list(na.rm = TRUE, trim = 0.1)
exec("mean", x = 1:10, !!!params)
```

```{r}
arg_name <- "na.rm"
arg_val <- TRUE
exec("mean", 1:10, !!arg_name := arg_val)
```

```{r}
x <- c(runif(10), NA)
funs <- c("mean", "median", "sd")

purrr::map_dbl(funs, exec, x, na.rm = TRUE)
```

### 19.6.3 dots_list()

```{r}
# Can easily move x to first entry:
tibble::tibble(
  y = 1:5,
  z = 3:-1,
  x = 5:1,
)

# Need to remove comma from z and add comma to x
data.frame(
  y = 1:5,
  z = 3:-1,
  x = 5:1
)
```

```{r, error = TRUE}
str(dots_list(x = 1, x = 2))


str(dots_list(x = 1, x = 2, .homonyms = "first"))


str(dots_list(x = 1, x = 2, .homonyms = "last"))


str(dots_list(x = 1, x = 2, .homonyms = "error"))
```

### 19.6.4 With base R

```{r}
do.call("rbind", dfs)
```

```{r}
args <- list(val)
names(args) <- var

do.call("data.frame", args)
```

```{r}
f <- function(...) {
  dots <- list(...)
  if (length(dots) == 1 && is.list(dots[[1]])) {
    dots <- dots[[1]]
  }
  
  # Do something
  ...
}
```

```{r}
f <- function(..., .dots) {
  dots <- c(list(...), .dots)
  # Do something
}
```

### 19.6.5 Exercises

1. One way to implement exec() is shown below. Describe how it works. What are the key ideas?

```{r}
exec <- function(f, ..., .env = caller_env()) {
  args <- list2(...)
  do.call(f, args, envir = .env)
}
```

This new exec command works the same as the previous example except it allows you to specify the environment from which to search for variables from. It then runs the function and along with the provided args. By default it uses the environment of the exec function

2. Carefully read the source code for `interaction()`, `expand.grid()`, and `par()`. Compare and contrast the techniques they use for switching between dots and list behaviour.

```{r}
#interaction
#par
#expand.grid
```
`interaction` first gathers ... into an object called args using `list`, checks to make sure it's not empty. If ... is of length 1 and is a list it changes its args and narg arguments. It then processes each item in args.

`par` first gathers ... into an object called args using `list`. It checks if args is empty and adds .Pars if it is

`expand.grid` does the same thing and decides what to do based on the length of nargs

3. Explain the problem with this definition of `set_attr()`

```{r, error = TRUE}
set_attr <- function(x, ...) {
  attr <- rlang::list2(...)
  attributes(x) <- attr
  x
}
set_attr(1:10, x = 10)

set_attr <- function(.x, ...) {
  attr <- rlang::list2(...)
  attributes(.x) <- attr
  .x
}
set_attr(1:10, x = 10)
```

The use of `x` instead of `.x` leads to name conflict issues

## 19.7 Case studies

### 19.7.1 lobstr::ast()
```{r}
z <- expr(foo(x, y))
lobstr::ast(z)
```

```{r}
lobstr::ast(!!z)
```

### 19.7.2 Map-reduce to generate code

```{r}
intercept <- 10
coefs <- c(x1 = 5, x2 = -4)
```

```{r}
coef_sym <- syms(names(coefs))
coef_sym
```

```{r}
summands <- map2(coef_sym, coefs, ~ expr((!!.x * !!.y)))
summands
```

```{r}
summands <- c(intercept, summands)
summands
```

```{r}
eq <- reduce(summands, ~ expr(!!.x + !!.y))
eq
```

```{r}
var <- expr(y)
coef_sym <- map(seq_along(coefs), ~ expr((!!var)[[!!.x]]))
coef_sym
```

```{r}
linear <- function(var, val) {
  var <- ensym(var)
  coef_name <- map(seq_along(val[-1]), ~ expr((!!var)[[!!.x]]))

  summands <- map2(val[-1], coef_name, ~ expr((!!.x * !!.y)))
  summands <- c(val[[1]], summands)

  reduce(summands, ~ expr(!!.x + !!.y))
}

linear(x, c(10, 5, -4))
```

### 19.7.3 Slicing an array

```{r}
indices <- rep(list(missing_arg()), 3)
expr(x[!!!indices])
```

```{r}
indices[[2]] <- 1
expr(x[!!!indices])
```

```{r}
slice <- function(x, along, index) {
  stopifnot(length(along) == 1)
  stopifnot(length(index) == 1)
    
  nd <- length(dim(x))
  indices <- rep(list(missing_arg()), nd)
  indices[[along]] <- index
  
  expr(x[!!!indices])
}

x <- array(sample(30), c(5, 2, 3))

slice(x, 1, 3)

slice(x, 2, 2)

slice(x, 3, 1)
```

### 19.7.4 Creating functions

```{r}
new_function(
  exprs(x = , y = ), 
  expr({x + y})
)
```

```{r}
power <- function(exponent) {
  new_function(
    exprs(x = ), 
    expr({
      x ^ !!exponent
    }), 
    caller_env()
  )
}
power(0.5)
```

```{r}
curve(sin(exp(4 * x)), n = 1000)
```

```{r}
curve2 <- function(expr, xlim = c(0, 1), n = 100) {
  expr <- enexpr(expr)
  f <- new_function(exprs(x = ), expr)
  
  x <- seq(xlim[1], xlim[2], length = n)
  y <- f(x)

  plot(x, y, type = "l", ylab = expr_text(expr))
}
curve2(sin(exp(4 * x)), n = 1000)
```

### 19.7.5 Exercises

1. In the linear-model example, we could replace the `expr()` in `reduce(summands, ~ expr(!!.x + !!.y))` with `call2()`: `reduce(summands, call2, "+")`. Compare and contrast the two approaches. Which do you think is easier to read?

```{r, error = TRUE}
# expr version
linear <- function(var, val) {
  var <- ensym(var)
  coef_name <- map(seq_along(val[-1]), ~ expr((!!var)[[!!.x]]))

  summands <- map2(val[-1], coef_name, ~ expr((!!.x * !!.y)))
  summands <- c(val[[1]], summands)

  reduce(summands, ~ expr(!!.x + !!.y))
}

linear(x, c(10, 5, -4))

# call2 version
linear <- function(var, val) {
  var <- ensym(var)
  coef_name <- map(seq_along(val[-1]), ~ expr((!!var)[[!!.x]]))

  summands <- map2(val[-1], coef_name, ~ expr((!!.x * !!.y)))
  summands <- c(val[[1]], summands)

  reduce(summands, call2, "+")
}

linear(x, c(10, 5, -4))
```

The `expr` command makes it a lot obvious what is going on. Also the `call2` code isn't working

2. Re-implement the Box-Cox transform defined below using unquoting and `new_function()`:

```{r}
bc <- function(lambda) {
  new_function(
    exprs(x = ), 
    if (lambda == 0) {
      expr({log(x)})}
    else {
      expr({(x ^ lambda - 1) / lambda})
    },
    caller_env()
  )
}

bc(lambda = 0)

bc(lambda = 5)
```

3. Re-implement the simple `compose()` defined below using quasiquotation and `new_function()`:

```{r}
compose <- function(f, g) {
  f <- enexpr(f)
  g <- enexpr(g)
  new_function(
    exprs(... =),
    expr((!!f)((!!g)(...))),
    caller_env()
  )
}

compose(abs, sum)
wtf <- compose(abs, sum)
wtf(c(1,3,5,6))
wtf(c(1,3,5,6,-20))
```
## 19.8 History