---
title: "20 Evaluation"
output: html_document
date: "2023-07-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 20.1 Introduction

```{r}
library(rlang)
library(purrr)
```

## 20.2 Evaluation basics

```{r}
x <- 10
eval(expr(x))

y <- 2
eval(expr(x + y))
```

```{r}
eval(expr(x + y), env(x = 1000))
```

```{r}
eval(print(x + 1), env(x = 1000))
```

```{r}
eval(expr(print(x + 1)), env(x = 1000))
```

### 20.2.1 Application: `local()`

```{r, error = T}
# Clean up variables created earlier
rm(x, y)

foo <- local({
  x <- 10
  y <- 200
  x + y
})

foo

x

y
```

```{r, error = TRUE}
local2 <- function(expr) {
  env <- env(caller_env())
  eval(enexpr(expr), env)
}

foo <- local2({
  x <- 10
  y <- 200
  x + y
})

foo

x

y
```

### 20.2.2 Application: `source()`

```{r}
source2 <- function(path, env = caller_env()) {
  file <- paste(readLines(path, warn = FALSE), collapse = "\n")
  exprs <- parse_exprs(file)

  res <- NULL
  for (i in seq_along(exprs)) {
    res <- eval(exprs[[i]], env)
  }

  invisible(res)
}
```

```{r}
source3 <- function(file, env = parent.frame()) {
  lines <- parse(file)
  res <- eval(lines, envir = env)
  invisible(res)
}
```

### 20.2.3 Gotcha: `function()`

```{r}
x <- 10
y <- 20
f <- eval(expr(function(x, y) !!x + !!y))
f
```

```{r}
lobstr::ast(f)
```


```{r}
f()
```

```{r}
attr(f, "srcref") <- NULL
f
```


```{r}
lobstr::ast(f)
```
### 20.2.4 Exercises

1. Carefully read the documentation for `source()`. What environment does it use by default? What if you supply `local = TRUE`? How do you provide a custom environment?

Uses global with local = FALSE. With TRUE is uses the environment in which source is called. Or you can provide an environment to local instead of T or F

2. Predict the results of the following lines of code:

```{r}
eval(expr(eval(expr(eval(expr(2 + 2))))))
# 4
eval(eval(expr(eval(expr(eval(expr(2 + 2)))))))
# 4
expr(eval(expr(eval(expr(eval(expr(2 + 2)))))))
# Since outer most call is expr
# eval(expr(eval(expr(eval(expr(2 + 2)))))))
```

3. Fill in the function bodies below to re-implement `get()` using `sym()` and `eval()`, and `assign()` using `sym()`, `expr()`, and `eval()`. Don’t worry about the multiple ways of choosing an environment that `get()` and `assign()` support; assume that the user supplies it explicitly.

```{r}
new_get <- function(name, env) {
  eval(expr = sym(name), envir = env)
}
new_assign <- function(name, value, env){
  eval(expr = expr(!!sym(name) <- !!value), envir = env)
}

new_get("x", env = .GlobalEnv)

new_assign("a", value = 420, .GlobalEnv)
new_get("a", env = .GlobalEnv)
```


4. Modify `source2()` so it returns the result of every expression, not just the last one. Can you eliminate the for loop?

```{r}
source2 <- function(path, env = caller_env()) {
  file <- paste(readLines(path, warn = FALSE), collapse = "\n")
  exprs <- parse_exprs(file)

  res <- NULL
  for (i in seq_along(exprs)) {
    res <- eval(exprs[[i]], env)
  }

  invisible(res)
}

source2 <- function(path, env = caller_env()) {
  file <- paste(readLines(path, warn = FALSE), collapse = "\n")
  exprs <- parse_exprs(file)

  res <- map(exprs, eval, env)

  invisible(res)
}
```


5. We can make `base::local()` slightly easier to understand by spreading out over multiple lines:

```{r}
local3 <- function(expr, envir = new.env()) {
  call <- substitute(eval(quote(expr), envir))
  eval(call, envir = parent.frame())
}
```

Explain how `local()` works in words. (Hint: you might want to `print(call)` to help understand what `substitute()` is doing, and read the documentation to remind yourself what environment `new.env()` will inherit from.)

First a call to eval because it's inside of substitute, then it substitutes and then evals the new call in the caller environment

## 20.3 Quosures

### 20.3.1 Creating

```{r}
foo <- function(x) enquo(x)
foo(a + b)
```

```{r}
quo(x + y + z)
```

```{r}
new_quosure(expr(x + y), env(x = 1, y = 10))
```

### 20.3.2 Evaluating

```{r}
q1 <- new_quosure(expr(x + y), env(x = 1, y = 10))
eval_tidy(q1)
```

### 20.3.3 Dots

```{r}
f <- function(...) {
  x <- 1
  g(..., f = x)
}
g <- function(...) {
  enquos(...)
}

x <- 0
qs <- f(global = x)
qs
```

```{r}
map_dbl(qs, eval_tidy)
```

### 20.3.4 Under the hood

```{r}
f <- ~runif(3)
str(f)
```

```{r}
q4 <- new_quosure(expr(x + y + z))
class(q4)
```

```{r}
is_call(q4)

q4[[1]]

q4[[2]]
```

```{r}
attr(q4, ".Environment")
```

```{r}
get_expr(q4)

get_env(q4)
```

### 20.3.5 Nested quosures

```{r}
q2 <- new_quosure(expr(x), env(x = 1))
q3 <- new_quosure(expr(x), env(x = 10))

x <- expr(!!q2 + !!q3)
```

```{r}
eval_tidy(x)
```

```{r}
x
```

```{r}
expr_print(x)
```

### 20.3.6 Exercises

1. Predict what each of the following quosures will return if evaluated.

```{r}
q1 <- new_quosure(expr(x), env(x = 1))
q1
#> <quosure>
#> expr: ^x
#> env:  0x7fac62d19130

q2 <- new_quosure(expr(x + !!q1), env(x = 10))
q2
#> <quosure>
#> expr: ^x + (^x)
#> env:  0x7fac62e35a98

q3 <- new_quosure(expr(x + !!q2), env(x = 100))
q3
#> <quosure>
#> expr: ^x + (^x + (^x))
#> env:  0x7fac6302feb0
```

```{r}
eval_tidy(q1)
# 1
eval_tidy(q2)
#11
eval_tidy(q3)
#111
```
It's all nested shit

2. Write an `enenv()` function that captures the environment associated with an argument. (Hint: this should only require two function calls.)

```{r}
enenv <- function(x){
  get_env(enquo(x))
}

enenv(a)
```


## 20.4 Data masks
### 20.4.1 Basics

```{r}
q1 <- new_quosure(expr(x * y), env(x = 100))
df <- data.frame(y = 1:10)

eval_tidy(q1, df)
```

```{r}
with2 <- function(data, expr) {
  expr <- enquo(expr)
  eval_tidy(expr, data)
}
```

```{r}
x <- 100
with2(df, x * y)
```

```{r}
with3 <- function(data, expr) {
  expr <- substitute(expr)
  eval(expr, data, caller_env())
}
```

### 20.4.2 Pronouns

```{r}
with2(df, x)
```

```{r}
x <- 1
df <- data.frame(x = 2)

with2(df, .data$x)

with2(df, .env$x)
```

```{r, error=TRUE}
with2(df, .data$y)
```

### 20.4.3 Application: `subset()`

```{r}
sample_df <- data.frame(a = 1:5, b = 5:1, c = c(5, 3, 1, 4, 1))

# Shorthand for sample_df[sample_df$a >= 4, ]
subset(sample_df, a >= 4)


# Shorthand for sample_df[sample_df$b == sample_df$c, ]
subset(sample_df, b == c)
```

```{r}
subset2 <- function(data, rows) {
  rows <- enquo(rows)
  rows_val <- eval_tidy(rows, data)
  stopifnot(is.logical(rows_val))

  data[rows_val, , drop = FALSE]
}

subset2(sample_df, b == c)
```

### 20.4.4 Application: transform
```{r}
df <- data.frame(x = c(2, 3, 1), y = runif(3))
transform(df, x = -x, y2 = 2 * y)
```

```{r}
transform2 <- function(.data, ...) {
  dots <- enquos(...)

  for (i in seq_along(dots)) {
    name <- names(dots)[[i]]
    dot <- dots[[i]]

    .data[[name]] <- eval_tidy(dot, .data)
  }

  .data
}

transform2(df, x2 = x * 2, y = -y)
```

### 20.4.5 Application: `select()`

```{r}
df <- data.frame(a = 1, b = 2, c = 3, d = 4, e = 5)
subset(df, select = b:d)
```

```{r}
vars <- as.list(set_names(seq_along(df), names(df)))
str(vars)
```

```{r}
select2 <- function(data, ...) {
  dots <- enquos(...)

  vars <- as.list(set_names(seq_along(data), names(data)))
  cols <- unlist(map(dots, eval_tidy, vars))

  data[, cols, drop = FALSE]
}
select2(df, b:d)
```

### 20.4.6 Exercises
1. Why did I use a for loop in transform2() instead of map()? Consider `transform2(df, x = x * 2, x = x * 2)`.

```{r, error=TRUE}
df <- data.frame(x = c(2, 3, 1), y = runif(3))
transform2 <- function(.data, ...) {
  dots <- enquos(...)

  for (i in seq_along(dots)) {
    name <- names(dots)[[i]]
    dot <- dots[[i]]

    .data[[name]] <- eval_tidy(dot, .data)
  }

  .data
}

transform2(df, x2 = x * 2, y = -y)
transform2(df, x = x * 2, x = x * 2)
transform2(df, x2 = x * 2, x3 = x2 * 2)

transform3 <- function(.data, ...) {
  dots <- enquos(...)
  map(dots, eval_tidy, .data)
}

transform3(df, x2 = x * 2, y = -y)
transform3(df, x = x * 2, x = x * 2)
transform3(df, x2 = x * 2, x3 = x2 * 2)
```

Each operation is done in order. This allows your changes to x to be used in the next call which uses x in the chain of transformations. map doesn't allow for this since each is handled independently

2. Here’s an alternative implementation of `subset2()`:

Compare and contrast `subset3()` to `subset2()`. What are its advantages and disadvantages?


```{r}
#Orignal
subset2 <- function(data, rows) {
  rows <- enquo(rows)
  rows_val <- eval_tidy(rows, data)
  stopifnot(is.logical(rows_val))

  data[rows_val, , drop = FALSE]
}

#New
subset3 <- function(data, rows) {
  rows <- enquo(rows)
  eval_tidy(expr(data[!!rows, , drop = FALSE]), data = data)
}

df <- data.frame(x = 1:3)
subset2(df, x == 1)
subset3(df, x == 1)
```

Both work, definitely more error checking in the first function. Second does the subsetting in the context of the data mask since they are in the same call. Second is also harder to read. Can run into issues if there is a column called data.

3. The following function implements the basics of `dplyr::arrange()`. Annotate each line with a comment explaining what it does. Can you explain why !!.na.last is strictly correct, but omitting the !! is unlikely to cause problems?

```{r}
arrange2 <- function(.df, ..., .na.last = TRUE) {
  # get the arguments and order for ordering the rows
  args <- enquos(...) 
  # splice unquote the args into order along with how to treat NAs
  order_call <- expr(order(!!!args, na.last = !!.na.last))
  # Create the index for subsetting the data frame into the new order
  ord <- eval_tidy(order_call, .df)
  # Make sure the vector of indexes is the same length as the original data frame, meaning now rows were dropped.
  stopifnot(length(ord) == nrow(.df))
  # reorder the data frame using subsetting and the new ordering indices
  .df[ord, , drop = FALSE]
}
```



## 20.5 Using tidy evaluation

### 20.5.1 Quoting and unquoting
```{r}
resample <- function(df, n) {
  idx <- sample(nrow(df), n, replace = TRUE)
  df[idx, , drop = FALSE]
}
```

```{r, error=TRUE}
subsample <- function(df, cond, n = nrow(df)) {
  df <- subset2(df, cond)
  resample(df, n) # this n can
}
rm(x)
df <- data.frame(x = c(1, 1, 1, 2, 2), y = 1:5)
subsample(df, x == 1)
```

```{r}
subsample <- function(df, cond, n = nrow(df)) {
  cond <- enquo(cond)

  df <- subset2(df, !!cond)
  resample(df, n)
}

subsample(df, x == 1)
```

### 20.5.2 Handling ambiguity

```{r}
threshold_x <- function(df, val) {
  subset2(df, x >= val)
}
```

```{r}
x <- 10
no_x <- data.frame(y = 1:3)
threshold_x(no_x, 2)
```

```{r}
has_val <- data.frame(x = 1:3, val = 9:11)
threshold_x(has_val, 2)
```

```{r, error = TRUE}
threshold_x <- function(df, val) {
  subset2(df, .data$x >= .env$val)
}

x <- 10
threshold_x(no_x, 2)
#> Error: Column `x` not found in `.data`
threshold_x(has_val, 2)
```

```{r}
threshold_x <- function(df, val) {
  subset2(df, .data$x >= !!val)
}
```

### 20.5.3 Quoting and ambiguity

```{r}
threshold_var <- function(df, var, val) {
  var <- as_string(ensym(var))
  subset2(df, .data[[var]] >= !!val)
}

df <- data.frame(x = 1:10)
threshold_var(df, x, 8)
```

```{r}
threshold_expr <- function(df, expr, val) {
  expr <- enquo(expr)
  subset2(df, !!expr >= !!val)
}
```

### 20.5.4 Exercises

1. I’ve included an alternative implementation of `threshold_var()` below. What makes it different to the approach I used above? What makes it harder?

```{r}
# Old
threshold_var <- function(df, var, val) {
  var <- as_string(ensym(var))
  subset2(df, .data[[var]] >= !!val)
}

# New
threshold_var <- function(df, var, val) {
  var <- ensym(var)
  subset2(df, `$`(.data, !!var) >= !!val)
}
```


var no longer is wrapped with "as_string". Subsetting is now done using `$` in the prefix form. It overall is uglier to look at

## 20.6 Base evaluation
### 20.6.1 `substitute()`

```{r}
subset_base <- function(data, rows) {
  rows <- substitute(rows)
  rows_val <- eval(rows, data, caller_env())
  stopifnot(is.logical(rows_val))

  data[rows_val, , drop = FALSE]
}

subset_tidy <- function(data, rows) {
  rows <- enquo(rows)
  rows_val <- eval_tidy(rows, data)
  stopifnot(is.logical(rows_val))

  data[rows_val, , drop = FALSE]
}
```

#### 20.6.1.1 Programming with `subset()`

```{r}
f1 <- function(df, ...) {
  xval <- 3
  subset_base(df, ...)
}

my_df <- data.frame(x = 1:3, y = 3:1)
xval <- 1
f1(my_df, x == xval)
```

```{r, error = TRUE}
local({
  zzz <- 2
  dfs <- list(data.frame(x = 1:3), data.frame(x = 4:6))
  lapply(dfs, subset_base, x == zzz)
})
```

```{r}
f2 <- function(df1, expr) {
  call <- substitute(subset_base(df1, expr))
  expr_print(call)
  eval(call, caller_env())
}

my_df <- data.frame(x = 1:3, y = 3:1)
f2(my_df, x == 1)
```

```{r}
f3 <- function(df) {
  call <- substitute(subset_base(df, z > 0))
  expr_print(call)
  eval(call, caller_env())
}

my_df <- data.frame(x = 1:3, y = 3:1)
z <- -1
f3(my_df)
```


#### 20.6.1.2 What about `[`?

Need to set drop = FALSE, and remove NAs if using ]

### 20.6.2 `match.call()`

```{r}
g <- function(x, y, z) {
  match.call()
}
g(1, 2, z = 3)
```

```{r, eval = FALSE}
write.csv <- function(...) {
  call <- match.call(write.table, expand.dots = TRUE)

  call[[1]] <- quote(write.table)
  call$sep <- ","
  call$dec <- "."

  eval(call, parent.frame())
}
```

```{r, eval = FALSE}
write.csv <- function(...) {
  write.table(..., sep = ",", dec = ".")
}
```

#### 20.6.2.1 Wrapping modelling functions

```{r}
lm2 <- function(formula, data) {
  lm(formula, data)
}
```

```{r}
lm2(mpg ~ disp, mtcars)
```

```{r}
lm3 <- function(formula, data, env = caller_env()) {
  formula <- enexpr(formula)
  data <- enexpr(data)

  lm_call <- expr(lm(!!formula, data = !!data))
  expr_print(lm_call)
  eval(lm_call, env)
}

lm3(mpg ~ disp, mtcars)
```

```{r}
resp <- expr(mpg)
disp1 <- expr(vs)
disp2 <- expr(wt)
lm3(!!resp ~ !!disp1 + !!disp2, mtcars)
```

#### 20.6.2.2 Evaluation environment

```{r, error = TRUE}
resample_lm0 <- function(formula, data, env = caller_env()) {
  formula <- enexpr(formula)
  resample_data <- resample(data, n = nrow(data))

  lm_call <- expr(lm(!!formula, data = resample_data))
  expr_print(lm_call)
  eval(lm_call, env)
}

df <- data.frame(x = 1:10, y = 5 + 3 * (1:10) + round(rnorm(10), 2))
resample_lm0(y ~ x, data = df)
```

```{r}
resample_lm1 <- function(formula, data, env = caller_env()) {
  formula <- enexpr(formula)
  resample_data <- resample(data, n = nrow(data))

  lm_call <- expr(lm(!!formula, data = !!resample_data))
  expr_print(lm_call)
  eval(lm_call, env)
}
resample_lm1(y ~ x, data = df)$call
```

```{r}
resample_lm2 <- function(formula, data, env = caller_env()) {
  formula <- enexpr(formula)
  resample_data <- resample(data, n = nrow(data))

  lm_env <- env(env, resample_data = resample_data)
  lm_call <- expr(lm(!!formula, data = resample_data))
  expr_print(lm_call)
  eval(lm_call, lm_env)
}
resample_lm2(y ~ x, data = df)
```

### 20.6.3 Exercises

1. Why does this function fail?

```{r, error = TRUE}
lm3a <- function(formula, data) {
  formula <- enexpr(formula)

  lm_call <- expr(lm(!!formula, data = data))
  eval(lm_call, caller_env())
}
lm3a(mpg ~ disp, mtcars)$call

#Need to unquote data in the expr call

lm3a <- function(formula, data) {
  formula <- enexpr(formula)

  lm_call <- expr(lm(!!formula, data = !!data))
  eval(lm_call, caller_env())
}
lm3a(mpg ~ disp, mtcars)$call
```

2. When model building, typically the response and data are relatively constant while you rapidly experiment with different predictors. Write a small wrapper that allows you to reduce duplication in the code below.

```{r}
lm(mpg ~ disp, data = mtcars)
lm(mpg ~ I(1 / disp), data = mtcars)
lm(mpg ~ disp * cyl, data = mtcars)

new_lm <- function(predictors, env = caller_env()){
  predictors <- enexpr(predictors)
  data <- enexpr(data)
  
  lm_call <- expr(lm(mpg ~ !!predictors, data = mtcars))
  expr_print(lm_call)
  eval(lm_call, env)
}

lm(mpg ~ disp, data = mtcars)
new_lm(disp)

lm(mpg ~ I(1 / disp), data = mtcars)
new_lm(I(1 / disp))

lm(mpg ~ disp * cyl, data = mtcars)
new_lm(disp * cyl)
```

3. Another way to write `resample_lm()` would be to include the resample expression (`data[sample(nrow(data), replace = TRUE), , drop = FALSE]`) in the data argument. Implement that approach. What are the advantages? What are the disadvantages?

```{r}
df <- data.frame(x = 1:10, y = 5 + 3 * (1:10) + round(rnorm(10), 2))
resample_lm2(y ~ x, data = df)
resample_lm2(y ~ x, data = df[sample(nrow(df), replace = TRUE), , drop = FALSE])
```

Advantage: No need to create a new environment for the new data frame or the need to unquote it

Disadvantage: It's quite cumbersome for multiple calls. It also is hard coded for each call. Capturing becomes worse.