---
title: "25 Rewriting R code in C++"
output: html_document
date: "2023-09-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 25 Rewriting R code in C++
## 25.1 Introduction

```{r}
library(Rcpp)
```

## 25.2 Getting started with C++

```{r}
cppFunction('int add(int x, int y, int z) {
  int sum = x + y + z;
  return sum;
}')

add

add(1, 2, 3)
```

### 25.2.1 No inputs, scalar output

```{r}
one <- function() 1L
```

```
int one() {
  return 1;
}
```

```{r}
cppFunction('int one() {
  return 1;
}')
```

### 25.2.2 Scalar input, scalar output

```{r}
signR <- function(x) {
  if (x > 0) {
    1
  } else if (x == 0) {
    0
  } else {
    -1
  }
}

cppFunction('int signC(int x) {
  if (x > 0) {
    return 1;
  } else if (x == 0) {
    return 0;
  } else {
    return -1;
  }
}')
```

### 25.2.3 Vector input, scalar output

```{r}
sumR <- function(x) {
  total <- 0
  for (i in seq_along(x)) {
    total <- total + x[i]
  }
  total
}
```

```{r}
cppFunction('double sumC(NumericVector x) {
  int n = x.size();
  double total = 0;
  for(int i = 0; i < n; ++i) {
    total += x[i];
  }
  return total;
}')
```

```{r}
x <- runif(1e3)
bench::mark(
  sum(x),
  sumC(x),
  sumR(x)
)[1:6]
```

### 25.2.4 Vector input, vector output

```{r}
pdistR <- function(x, ys) {
  sqrt((x - ys) ^ 2)
}
```


```{r}
cppFunction('NumericVector pdistC(double x, NumericVector ys) {
  int n = ys.size();
  NumericVector out(n);

  for(int i = 0; i < n; ++i) {
    out[i] = sqrt(pow(ys[i] - x, 2.0));
  }
  return out;
}')
```

```{r}
y <- runif(1e6)
bench::mark(
  pdistR(0.5, y),
  pdistC(0.5, y)
)[1:6]
```

### 25.2.5 Using sourceCpp

```
#include <Rcpp.h>
using namespace Rcpp;
```

```
// [[Rcpp::export]]
```

```
/*** R
# This is R code
*/
```

```{r}
sourceCpp("cpp_example.cpp")
```

### 25.2.6 Exercises

1. With the basics of C++ in hand, it’s now a great time to practice by reading and writing some simple C++ functions. For each of the following functions, read the code and figure out what the corresponding base R function is. You might not understand every part of the code yet, but you should be able to figure out the basics of what the function does.

```{Rcpp}
#include <Rcpp.h>
using namespace Rcpp;

// [[Rcpp::export]]
double f1(NumericVector x) {
  int n = x.size();
  double y = 0;

  for(int i = 0; i < n; ++i) {
    y += x[i] / n;
  }
  return y;
}
```

```{r}
x <- c(1,2,3)
mean(x)
f1(x)
```
#### mean()

```{Rcpp}
#include <Rcpp.h>
using namespace Rcpp;

// [[Rcpp::export]]
NumericVector f2(NumericVector x) {
  int n = x.size();
  NumericVector out(n);

  out[0] = x[0];
  for(int i = 1; i < n; ++i) {
    out[i] = out[i - 1] + x[i];
  }
  return out;
}
```

```{r}
x <- c(1,2,3,4)
cumsum(x)
f2(x)
```
#### cumsum()

```{Rcpp}
#include <Rcpp.h>
using namespace Rcpp;

// [[Rcpp::export]]
bool f3(LogicalVector x) {
  int n = x.size();

  for(int i = 0; i < n; ++i) {
    if (x[i]) return true;
  }
  return false;
}
```

```{r}
x <- c(T,T,F)
any(x)
f3(x)
```
#### any()

```{Rcpp}
#include <Rcpp.h>
using namespace Rcpp;

// [[Rcpp::export]]
int f4(Function pred, List x) {
  int n = x.size();

  for(int i = 0; i < n; ++i) {
    LogicalVector res = pred(x[i]);
    if (res[0]) return i + 1;
  }
  return 0;
}
```

```{r}
f4(sample, c(1,2))
Position(sample, c(1,2))
```
#### Position()

```{Rcpp}
#include <Rcpp.h>
using namespace Rcpp;

// [[Rcpp::export]]
NumericVector f5(NumericVector x, NumericVector y) {
  int n = std::max(x.size(), y.size());
  NumericVector x1 = rep_len(x, n);
  NumericVector y1 = rep_len(y, n);

  NumericVector out(n);

  for (int i = 0; i < n; ++i) {
    out[i] = std::min(x1[i], y1[i]);
  }

  return out;
}
```

```{r}
x <- c(1,2,3,4,5)
y <- c(11,12,13,14,15)
f5(x,y)
pmin(x,y)
```
#### pmin()

2. To practice your function writing skills, convert the following functions into C++. For now, assume the inputs have no missing values.

`all()`
```{Rcpp}
#include <Rcpp.h>
using namespace Rcpp;

// [[Rcpp::export]]
bool allC(LogicalVector x) {
  int n = x.size();
  
  for(int i = 0; i < n; i++){
    if(!x[i]) return false;
  }
  return true;
}
```

```{r}
x <- c(T,T,F)
y <- c(T,T,T)

all(x)
allC(x)
all(y)
allC(y)
```

`cumprod()`
```{Rcpp}
#include <Rcpp.h>
using namespace Rcpp;

// [[Rcpp::export]]
NumericVector cumprodC(NumericVector x) {
  int n = x.size();
  NumericVector out(n);

  out[0] = x[0];
  for(int i = 1; i < n; ++i) {
    out[i] = out[i - 1] * x[i];
  }
  return out;
}
```

```{r}
x <- c(1,2,3,4)
cumprod(x)
cumprodC(x)
```

`cummin()`
```{Rcpp}
#include <Rcpp.h>
using namespace Rcpp;

// [[Rcpp::export]]
NumericVector cumminC(NumericVector x) {
  int n = x.size();
  NumericVector out(n);

  out[0] = x[0];
  for(int i = 1; i < n; ++i) {
      out[i] = std::min(x[i], out[i-1]);
  }
  return out;
}
```

```{r}
x <- c(6432,3462,234,6,1325235,1231,23525,2)
cummin(x)
cumminC(x)
```


`cummax()`
```{Rcpp}
#include <Rcpp.h>
using namespace Rcpp;

// [[Rcpp::export]]
NumericVector cummaxC(NumericVector x) {
  int n = x.size();
  NumericVector out(n);

  out[0] = x[0];
  for(int i = 1; i < n; ++i) {
      out[i] = std::max(x[i], out[i-1]);
  }
  return out;
}
```

```{r}
x <- c(6432,3462,234,6,1325235,1231,23525,2)
cummax(x)
cummaxC(x)
```

`diff()`
```{Rcpp}
#include <Rcpp.h>
using namespace Rcpp;

// [[Rcpp::export]]
NumericVector diffC(NumericVector x, int y) {
  int n = x.size() - y;
  NumericVector out(n);

  for(int i = 0; i < n; ++i) {
      out[i] = x[i + y] - x[i];
  }
  return out;
}


```

```{r}
x <- c(10,5,1,5,2,6)
diff(x,1)
diffC(x,1)
diff(x,2)
diffC(x,2)
```


`range()`
```{Rcpp}
#include <Rcpp.h>
using namespace Rcpp;

// [[Rcpp::export]]
NumericVector rangeC(NumericVector x) {
  int n = x.size();
  NumericVector out(2);
  
  out[0] = x[0];
  out[1] = x[0];
  
  for(int i = 0; i < n; ++i) {
      if(x[i] < out[0]) out[0] = x[i];
      if(x[i] > out[1]) out[1] = x[i];
  }
  
  return out;
}
```

```{r}
x <- c(234,12,63)
range(x)
rangeC(x)
```

`var()`
```{Rcpp}
#include <Rcpp.h>
using namespace Rcpp;

// [[Rcpp::export]]
double varC(NumericVector x){
  int n = x.size();
  double meanx = 0;
  
  for(int i = 0; i < n; i++){
    meanx += x[i] / n;
    }

  double var = 0;
  
  for(int i = 0; i < n; i++){
    var += pow(x[i] - meanx, 2);
    }
  return var / (n -1 );
}
```

```{r}
var(1:10)
varC(1:10)
```

## 25.3 Other classes

### 25.3.1 Lists and data frames

```{Rcpp}
#include <Rcpp.h>
using namespace Rcpp;

// [[Rcpp::export]]
double mpe(List mod) {
  if (!mod.inherits("lm")) stop("Input must be a linear model");

  NumericVector resid = as<NumericVector>(mod["residuals"]);
  NumericVector fitted = as<NumericVector>(mod["fitted.values"]);

  int n = resid.size();
  double err = 0;
  for(int i = 0; i < n; ++i) {
    err += resid[i] / (fitted[i] + resid[i]);
  }
  return err / n;
}
```


```{r}
mod <- lm(mpg ~ wt, data = mtcars)
mpe(mod)
```

### 25.3.2 Functions

```{Rcpp}
#include <Rcpp.h>
using namespace Rcpp;

// [[Rcpp::export]]
RObject callWithOne(Function f) {
  return f(1);
}
```

```{r}
callWithOne(function(x) x + 1)

callWithOne(paste)
```

```{r, eval = F}
f("y", 1);
```

```{r, eval = F}
f(_["x"] = "y", _["value"] = 1);
```

### 25.3.3 Attributes

```{Rcpp}
#include <Rcpp.h>
using namespace Rcpp;

// [[Rcpp::export]]
NumericVector attribs() {
  NumericVector out = NumericVector::create(1, 2, 3);

  out.names() = CharacterVector::create("a", "b", "c");
  out.attr("my-attr") = "my-value";
  out.attr("class") = "my-class";

  return out;
}
```

## 25.4 Missing values

### 25.4.1 Scalars

```{Rcpp}
#include <Rcpp.h>
using namespace Rcpp;

// [[Rcpp::export]]
List scalar_missings() {
  int int_s = NA_INTEGER;
  String chr_s = NA_STRING;
  bool lgl_s = NA_LOGICAL;
  double num_s = NA_REAL;

  return List::create(int_s, chr_s, lgl_s, num_s);
}
```

```{r}
str(scalar_missings())
```

#### 25.4.1.1 Integers

#### 25.4.1.2 Doubles

```{r}
evalCpp("NAN == 1")

evalCpp("NAN < 1")

evalCpp("NAN > 1")

evalCpp("NAN == NAN")
```

```{r}
evalCpp("NAN && TRUE")

evalCpp("NAN || FALSE")
```

```{r}
evalCpp("NAN + 1")

evalCpp("NAN - 1")

evalCpp("NAN / 1")

evalCpp("NAN * 1")
```

### 25.4.2 Strings
### 25.4.3 Boolean
### 25.4.4 Vectors

```{Rcpp}
#include <Rcpp.h>
using namespace Rcpp;

// [[Rcpp::export]]
List missing_sampler() {
  return List::create(
    NumericVector::create(NA_REAL),
    IntegerVector::create(NA_INTEGER),
    LogicalVector::create(NA_LOGICAL),
    CharacterVector::create(NA_STRING)
  );
}
```

```{r}
str(missing_sampler())
```

### 25.4.5 Exercises

1. Rewrite any of the functions from the first exercise of Section 25.2.6 to deal with missing values. If `na.rm` is true, ignore the missing values. If `na.rm` is false, return a missing value if the input contains any missing values. Some good functions to practice with are `min()`, `max()`, `range()`, `mean()`, and `var()`.

`mean()`
```{Rcpp}
#include <Rcpp.h>
using namespace Rcpp;

// [[Rcpp::export]]
double f1(NumericVector x, bool na_rm = false) {
  int og_n = x.size();
  int n = x.size();
  double y = 0;
  
  if (na_rm){
    for(int i = 0; i < og_n; ++i) {
      if (NumericVector::is_na(x[i])) {
        n -= 1;
        continue;
        }
      y += x[i];
      }
  } else {
    for(int i = 0; i < n; ++i) {
      if (NumericVector::is_na(x[i])) {
        return NA_REAL;
        }
      y += x[i];
      }
}
  return y/n;
}
```

```{r}
x <- c(1,2,3,4, NA, 5, 64, 123 ,NA, 123)
mean(x)
f1(x, na_rm = F)

mean(x, na.rm = T)
f1(x, na_rm = TRUE)
```

2. Rewrite `cumsum()` and `diff()` so they can handle missing values. Note that these functions have slightly more complicated behaviour.

`cumsum()`
```{Rcpp}
#include <Rcpp.h>
using namespace Rcpp;

// [[Rcpp::export]]
NumericVector f2(NumericVector x, bool na_rm = true) {
  int n = x.size();
  NumericVector out(n);
  out[0] = x[0];
  
  if (na_rm) {
    for (int i = 1; i < n; ++i) {
      if (NumericVector::is_na(x[i])) {
        out[i] = out[i - 1];
      } else {
        out[i] = out[i - 1] + x[i];
        }
      }
    } else {
      for(int i = 1; i < n; ++i) {
        if (NumericVector::is_na(x[i])) {
          out[i] = NA_REAL;
        } else {
          out[i] = out[i - 1] + x[i];
          }
        }
      }
    return out;
    }
```

```{r}
x <- c(1,2,3,4, NA, 5, 11)
cumsum(x)
f2(x)
f2(x, na_rm = FALSE)
```

Alternate`cumsum()`
```{Rcpp}
#include <Rcpp.h>
using namespace Rcpp;

// [[Rcpp::export]]
NumericVector alt_f2(NumericVector x, bool na_rm = true) {
  int og_n = x.size();
  int n = x.size();
  int skips = 0;

  if (na_rm) {
    for (int i = 1; i < og_n; ++i) { 
      if (NumericVector::is_na(x[i])) {
        n -= 1;
        }
      }
  }
  
  NumericVector out(n);
  out[0] = x[0];
  
  if (na_rm) {
    for (int i = 1; i < og_n; ++i) {
      if (NumericVector::is_na(x[i])) {
        skips += 1;
        continue;
      } else {
        out[i - skips] = out[i - 1 - skips] + x[i];
        }
    }
  } else {
    for(int i = 1; i < og_n; ++i) {
      if (NumericVector::is_na(x[i])) {
        out[i] = NA_REAL;
      } else {
        out[i] = out[i - 1] + x[i];
        }
      }
    }
  return out;
  }
```

```{r}
x <- c(1,2,3,4, NA, 5, 11)
cumsum(x)
alt_f2(x, na_rm = FALSE)
alt_f2(x)
```

`diff()`
```{Rcpp}
#include <Rcpp.h>
using namespace Rcpp;

// [[Rcpp::export]]
NumericVector diffC(NumericVector x, int y) {
  int n = x.size() - y;
  NumericVector out(n);

  for(int i = 0; i < n; ++i) {
      out[i] = x[i + y] - x[i];
  }
  return out;
}


```

```{r}
x <- c(10,5,NA,5,2,6)
diff(x,1)
diffC(x,1)
diff(x,2)
diffC(x,2)
```

Already Works?

## 25.5 Standard Template Library
### 25.5.1 Using iterators

```{Rcpp}
#include <Rcpp.h>
using namespace Rcpp;

// [[Rcpp::export]]
double sum3(NumericVector x) {
  double total = 0;
  
  NumericVector::iterator it;
  for(it = x.begin(); it != x.end(); ++it) {
    total += *it;
  }
  return total;
}
```

```{Rcpp}
// [[Rcpp::plugins(cpp11)]]
#include <Rcpp.h>
using namespace Rcpp;

// [[Rcpp::export]]
double sum4(NumericVector xs) {
  double total = 0;
  
  for(const auto &x : xs) {
    total += x;
  }
  return total;
}
```

```{Rcpp}
#include <numeric>
#include <Rcpp.h>
using namespace Rcpp;

// [[Rcpp::export]]
double sum5(NumericVector x) {
  return std::accumulate(x.begin(), x.end(), 0.0);
}
```

### 25.5.2 Algorithms

```{Rcpp}
#include <algorithm>
#include <Rcpp.h>
using namespace Rcpp;

// [[Rcpp::export]]
IntegerVector findInterval2(NumericVector x, NumericVector breaks) {
  IntegerVector out(x.size());

  NumericVector::iterator it, pos;
  IntegerVector::iterator out_it;

  for(it = x.begin(), out_it = out.begin(); it != x.end(); 
      ++it, ++out_it) {
    pos = std::upper_bound(breaks.begin(), breaks.end(), *it);
    *out_it = std::distance(breaks.begin(), pos);
  }

  return out;
}
```

#### 25.5.3 Data structures
#### 25.5.4 Vectors

```{Rcpp}
#include <Rcpp.h>
using namespace Rcpp;

// [[Rcpp::export]]
List rleC(NumericVector x) {
  std::vector<int> lengths;
  std::vector<double> values;

  // Initialise first value
  int i = 0;
  double prev = x[0];
  values.push_back(prev);
  lengths.push_back(1);

  NumericVector::iterator it;
  for(it = x.begin() + 1; it != x.end(); ++it) {
    if (prev == *it) {
      lengths[i]++;
    } else {
      values.push_back(*it);
      lengths.push_back(1);

      i++;
      prev = *it;
    }
  }

  return List::create(
    _["lengths"] = lengths, 
    _["values"] = values
  );
}
```

### 25.5.5 Sets

```{Rcpp}
// [[Rcpp::plugins(cpp11)]]
#include <Rcpp.h>
#include <unordered_set>
using namespace Rcpp;

// [[Rcpp::export]]
LogicalVector duplicatedC(IntegerVector x) {
  std::unordered_set<int> seen;
  int n = x.size();
  LogicalVector out(n);

  for (int i = 0; i < n; ++i) {
    out[i] = !seen.insert(x[i]).second;
  }

  return out;
}
```

### 25.5.6 Map

```{Rcpp}
#include <Rcpp.h>
using namespace Rcpp;

// [[Rcpp::export]]
std::map<double, int> tableC(NumericVector x) {
  std::map<double, int> counts;

  int n = x.size();
  for (int i = 0; i < n; i++) {
    counts[x[i]]++;
  }

  return counts;
}
```

### 25.5.7 Exercises

1. median.default() using partial_sort.

```{Rcpp}
#include <Rcpp.h>
#include <algorithm>
using namespace Rcpp;

// [[Rcpp::export]]
double medianC(NumericVector x) {
  int n = x.size();
  
  if (n % 2 == 0) {
    std::partial_sort (x.begin(), x.begin() + n / 2 + 1, x.end());
    return (x[n / 2 - 1] + x[n / 2]) / 2;
  } else {
    std::partial_sort (x.begin(), x.begin() + (n + 1) /2, x.end());
    return x[(n + 1) / 2 - 1];
    }
}

```

```{r}
x <- c(3,1,5,2,4)
y <- c(1,3,5,2,4,6)
median.default(x)
medianC(x)
median.default(y)
medianC(y)
```

2. %in% using unordered_set and the find() or count() methods.

```{Rcpp}
#include <Rcpp.h>
#include <algorithm>
using namespace Rcpp;

// [[Rcpp::export]]
LogicalVector inC(CharacterVector x, CharacterVector table) {
  std::unordered_set<String> seen;
  seen.insert(table.begin(), table.end());
  
  int n = x.size();
  LogicalVector out(n);
  for (int i = 0; i < n; ++i) {
    out[i] = seen.find(x[i]) != seen.end();
  }
  
  return out;
}

```

```{r}
x <- c(1,2,3)
c(3,4) %in% x
inC(c(3,4), x)
```

3. unique() using an unordered_set (challenge: do it in one line!).

```{Rcpp}
#include <Rcpp.h>
#include <unordered_set>
using namespace Rcpp;

// [[Rcpp::export]]
std::unordered_set<double> uniqueC(NumericVector x) {
  return std::unordered_set<double>(x.begin(), x.end());
}
```

```{r}
x <- c(1,1,34,12,534,123,34)
unique(x)
uniqueC(x)
```

4. min() using std::min(), or max() using std::max().

```{Rcpp}
#include <Rcpp.h>
#include <algorithm>
using namespace Rcpp;
// [[Rcpp::export]]
double maxC(NumericVector x) {
  double out = x[0];
  
  NumericVector::iterator it;
  for(it = x.begin(); it != x.end(); it++){
    out = std::max(out, *it);
    }
  return out;
}
```

```{r}
x <- c(1,1,34,12,534,123,34)
max(x)
maxC(x)
```
5. which.min() using min_element, or which.max() using max_element.

```{Rcpp}
#include <Rcpp.h>
#include <algorithm>
using namespace Rcpp;
// [[Rcpp::export]]
double which_maxC(NumericVector x) {
  int out = std::distance(x.begin(),
                          std::max_element(x.begin(),
                                           x.end())) + 1;
  return out;
}
```

```{r}
x <- c(1,1,34,12,534,123,34)
which.max(x)
which_maxC(x)
```



6. setdiff(), union(), and intersect() for integers using sorted ranges and set_union, set_intersection and set_difference.

```{Rcpp}
#include <Rcpp.h>
#include <algorithm>
using namespace Rcpp;
// [[Rcpp::export]]
IntegerVector setdiffC(IntegerVector x, IntegerVector y){
  IntegerVector sorted_x = clone(x);
  IntegerVector sorted_y = clone(y);
  IntegerVector out(x.size());
  IntegerVector::iterator it;
  
  std::sort(sorted_x.begin(), sorted_x.end());
  std::sort(sorted_y.begin(), sorted_y.end());
  
  std::set_difference(sorted_x.begin(), sorted_x.end(),
                           sorted_y.begin(), sorted_y.end(),
                           out.begin());
  
  return out;
}
```

```{r}
x <- c(1,2,3,4)
y <- c(3,4,5,6)
setdiff(x,y)
setdiffC(x,y)
```

## 25.6 Case studies

### 25.6.1 Gibbs sampler

```{r}
gibbs_r <- function(N, thin) {
  mat <- matrix(nrow = N, ncol = 2)
  x <- y <- 0

  for (i in 1:N) {
    for (j in 1:thin) {
      x <- rgamma(1, 3, y * y + 4)
      y <- rnorm(1, 1 / (x + 1), 1 / sqrt(2 * (x + 1)))
    }
    mat[i, ] <- c(x, y)
  }
  mat
}
```

```{Rcpp}
#include <Rcpp.h>
using namespace Rcpp;

// [[Rcpp::export]]
NumericMatrix gibbs_cpp(int N, int thin) {
  NumericMatrix mat(N, 2);
  double x = 0, y = 0;

  for(int i = 0; i < N; i++) {
    for(int j = 0; j < thin; j++) {
      x = rgamma(1, 3, 1 / (y * y + 4))[0];
      y = rnorm(1, 1 / (x + 1), 1 / sqrt(2 * (x + 1)))[0];
    }
    mat(i, 0) = x;
    mat(i, 1) = y;
  }

  return(mat);
}
```

```{r}
bench::mark(
  gibbs_r(100, 10),
  gibbs_cpp(100, 10),
  check = FALSE
)
```

### 25.6.2 R vectorisation versus C++ vectorisation

```{r}
vacc1a <- function(age, female, ily) {
  p <- 0.25 + 0.3 * 1 / (1 - exp(0.04 * age)) + 0.1 * ily
  p <- p * if (female) 1.25 else 0.75
  p <- max(0, p)
  p <- min(1, p)
  p
}
```

```{r}
vacc1 <- function(age, female, ily) {
  n <- length(age)
  out <- numeric(n)
  for (i in seq_len(n)) {
    out[i] <- vacc1a(age[i], female[i], ily[i])
  }
  out
}
```

```{r}
vacc2 <- function(age, female, ily) {
  p <- 0.25 + 0.3 * 1 / (1 - exp(0.04 * age)) + 0.1 * ily
  p <- p * ifelse(female, 1.25, 0.75)
  p <- pmax(0, p)
  p <- pmin(1, p)
  p
}
```

```{r}
vacc2a <- function(age, female, ily) {
  p <- 0.25 + 0.3 * 1 / (1 - exp(0.04 * age)) + 0.1 * ily
  p <- p * 0.75 + p * 0.5 * female
  p[p < 0] <- 0
  p[p > 1] <- 1
  p
}
```

```{Rcpp}
#include <Rcpp.h>
using namespace Rcpp;

double vacc3a(double age, bool female, bool ily){
  double p = 0.25 + 0.3 * 1 / (1 - exp(0.04 * age)) + 0.1 * ily;
  p = p * (female ? 1.25 : 0.75);
  p = std::max(p, 0.0);
  p = std::min(p, 1.0);
  return p;
}

// [[Rcpp::export]]
NumericVector vacc3(NumericVector age, LogicalVector female, 
                    LogicalVector ily) {
  int n = age.size();
  NumericVector out(n);

  for(int i = 0; i < n; ++i) {
    out[i] = vacc3a(age[i], female[i], ily[i]);
  }

  return out;
}
```

```{r}
n <- 1000
age <- rnorm(n, mean = 50, sd = 10)
female <- sample(c(T, F), n, rep = TRUE)
ily <- sample(c(T, F), n, prob = c(0.8, 0.2), rep = TRUE)

stopifnot(
  all.equal(vacc1(age, female, ily), vacc2(age, female, ily)),
  all.equal(vacc2(age, female, ily), vacc2a(age, female, ily)),
  all.equal(vacc1(age, female, ily), vacc3(age, female, ily))
)
```

```{r}
bench::mark(
  vacc1 = vacc1(age, female, ily),
  vacc2 = vacc2(age, female, ily),
  vacc2a = vacc2a(age, female, ily),
  vacc3 = vacc3(age, female, ily)
)
```

## 25.7 Using Rcpp in a package
## 25.8 Learning more
## 25.9 Acknowledgments