---
title: "6_Functions"
output: pdf_document
date: "2022-10-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

# 6. Functions

## 6.1 Introduction

### Quiz

1.  What are the three components of a function?

arguments, body, and environment

2.  What does the following code return?

```{r}
x <- 10
f1 <- function(x) {
  function() {
    x + 10
  }
}
f1(1)()
```

3.  How would you usually write this code?

```{r}
`+`(1, `*`(2, 3))
```

4.  How could you make this call easier to read?

```{r}
mean(, TRUE, x = c(1:10, NA))
```

5.  Does the following code throw an error when executed? Why or why not?

```{r}
f2 <- function(a, b) {
  a * 10
}
f2(10, stop("This is an error!"))
```

6.  What is an infix function? How do you write it? What's a replacement function? How do you write it?

7.  How do you ensure that cleanup action occurs regardless of how a function exits?

## 6.2 Function fundamentals

-   Functions can be broken down into three components: arguments, body, and environment.
-   Functions are objects, just as vectors are objects.

### 6.2.1 Function components

A function has three parts:

-   The formals(), the list of arguments that control how you call the function.
-   The body(), the code inside the function.
-   The environment(), the data structure that determines how the function finds the values associated with the names.

```{r}
f02 <- function(x, y) {
  # A comment
  x + y
}

formals(f02)

body(f02)

environment(f02)

```

```{r}
attr(f02, "srcref")
```

### 6.2.2 Primitive functions

```{r}
sum
`[`
```

```{r}
typeof(sum)
typeof(`[`)

```

```{r}
formals(sum)
body(sum)
environment(sum)
```

### 6.2.3 First-class functions

```{r}
f01 <- function(x) {
  sin(1 / x ^ 2)
}
```

```{r}
lapply(mtcars, function(x) length(unique(x)))
Filter(function(x) !is.numeric(x), mtcars)
integrate(function(x) sin(x) ^ 2, 0, pi)
```

```{r}
funs <- list(
  half = function(x) x / 2,
  double = function(x) x * 2
)

funs$double(10)
```

### 6.2.4 Invoking a function

```{r}
args <- list(1:10, na.rm = TRUE)
```

```{r}
do.call(mean, args)
```

#### 6.2.5 Exercises

1.  Given a name, like `"mean"`, [`match.fun()`](https://rdrr.io/r/base/match.fun.html) lets you find a function. Given a function, can you find its name? Why doesn't that make sense in R?

It's a lot easier to match a name then it is to match all the code in a function to a name

2.  It's possible (although typically not useful) to call an anonymous function. Which of the two approaches below is correct? Why?

```{r}
function(x) 3()
(function(x) 3)()
```

2nd?

3.  A good rule of thumb is that an anonymous function should fit on one line and shouldn't need to use {}. Review your code. Where could you have used an anonymous function instead of a named function? Where should you have used a named function instead of an anonymous function?

Should have used an anonymous on smaller quick functions. Use named functions on larget multi line chunks that need comments

4.  What function allows you to tell if an object is a function? What function allows you to tell if a function is a primitive function?

```{r}
is.primitive(sum)
is.function(mean)
```

5.  This code makes a list of all functions in the base package.

```{r}
objs <- mget(ls("package:base", all = TRUE), inherits = TRUE)
funs <- Filter(is.function, objs)
```

a.  Which base function has the most arguments?

```{r}
sort(unlist((lapply(
  lapply(funs, FUN = formals), length
))), decreasing = T) %>% head()
```

b.  How many base functions have no arguments? What's special about those functions?

```{r}
a <- unlist((lapply(
  lapply(funs, FUN = formals), length
)))
sum(a == 0)
head(a[a == 0])
is.primitive(`-`)
is.primitive("scan")
# Are primitive functions
```

c.  How could you adapt the code to find all primitive functions?

```{r}
objs <- mget(ls("package:base", all = TRUE), inherits = TRUE)
funs <- Filter(is.primitive, objs)
```

6.  What are the three important components of a function?

Formals, Body, Environment

7.  When does printing a function not show the environment it was created in?

Primitives and functions created in the global environment

## 6.3 Function composition

```{r}
square <- function(x) x^2
deviation <- function(x) x - mean(x)
```

```{r}
x <- runif(100)

sqrt(mean(square(deviation(x))))
```

```{r}
out <- deviation(x)
out <- square(out)
out <- mean(out)
out <- sqrt(out)
out
```

```{r}
library(magrittr)

x %>%
  deviation() %>%
  square() %>%
  mean() %>%
  sqrt()
```

## 6.4 Lexical scoping

```{r}
x <- 10
g01 <- function() {
  x <- 20
  x
}

g01()
```

R's lexical scoping follows four primary rules:

-   Name masking
-   Functions versus variables
-   A fresh start
-   Dynamic lookup

### 6.4.1 Name masking

```{r}
x <- 10
y <- 20
g02 <- function() {
  x <- 1
  y <- 2
  c(x, y)
}
g02()
```

```{r}
x <- 2
g03 <- function() {
  y <- 1
  c(x, y)
}
g03()
y
```
```{r}
x <- 1
g04 <- function() {
  y <- 2
  i <- function() {
    z <- 3
    c(x, y, z)
  }
  i()
}
g04()
```

### 6.4.2 Functions versus variables

```{r}
g07 <- function(x) x + 1
g08 <- function() {
  g07 <- function(x) x + 100
  g07(10)
}
g08()
```

```{r}
g09 <- function(x) x + 100
g10 <- function() {
  g09 <- 10
  g09(g09)
}
g10()
```

### 6.4.3 A fresh start

```{r}
g11 <- function() {
  if (!exists("a")) {
    a <- 1
  } else {
    a <- a + 1
  }
  a
}

g11()
g11()
```

### 6.4.4 Dynamic lookup

```{r}
g12 <- function() x + 1
x <- 15
g12()

x <- 20
g12()
```

```{r}
codetools::findGlobals(g12)
```

```{r, error = T}
environment(g12) <- emptyenv()
g12()
```

#### 6.4.5 Exercises

1. What does the following code return? Why? Describe how each of the three c’s is interpreted.

```{r}
c <- 10
c(c = c)
```
Creates a vector with one named element "c" with a value of 10

2. What are the four principles that govern how R looks for values?

Name masking, Functions versus variables, A fresh start, Dynamic lookup

3. What does the following function return? Make a prediction before running the code yourself.

```{r}
f <- function(x) {
  f <- function(x) {
    f <- function() {
      x ^ 2
    }
    f() + 1
  }
  f(x) * 2
}
f(10)
```

202

## 6.5 Lazy evaluation

```{r}
h01 <- function(x) {
  10
}
h01(stop("This is an error!"))
```

### 6.5.1 Promises

```{r}
y <- 10
h02 <- function(x) {
  y <- 100
  x + 1
}

h02(y)
```

```{r}
h02(y <- 1000)
y
```

```{r}
double <- function(x) { 
  message("Calculating...")
  x * 2
}

h03 <- function(x) {
  c(x, x)
}

h03(double(20))
```

### 6.5.2 Default arguments

```{r}
h04 <- function(x = 1, y = x * 2, z = a + b) {
  a <- 10
  b <- 100
  
  c(x, y, z)
}

h04()
```

```{r}
h05 <- function(x = ls()) {
  a <- 1
  x
}

# ls() evaluated inside h05:
h05()

# ls() evaluated in global environment:
h05(ls())
```

### 6.5.3 Missing arguments

```{r}
h06 <- function(x = 10) {
  list(missing(x), x)
}
str(h06())

str(h06(10))
```

```{r}
args(sample)
```

```{r, eval = F}
sample <- function(x, size = NULL, replace = FALSE, prob = NULL) {
  if (is.null(size)) {
    size <- length(x)
  }
  
  x[sample.int(length(x), size, replace = replace, prob = prob)]
}
```

```{r, eval = F}
`%||%` <- function(lhs, rhs) {
  if (!is.null(lhs)) {
    lhs
  } else {
    rhs
  }
}

sample <- function(x, size = NULL, replace = FALSE, prob = NULL) {
  size <- size %||% length(x)
  x[sample.int(length(x), size, replace = replace, prob = prob)]
}
```

#### 6.5.4 Exercises

1. What important property of && makes x_ok() work?

```{r}
x_ok <- function(x) {
  !is.null(x) && length(x) == 1 && x > 0
}

x_ok(NULL)
x_ok(1)
x_ok(1:3)
```

Processed from left to right. Prevents error later on by requiring the first tests to work.

```{r}
x_ok <- function(x) {
  !is.null(x) & length(x) == 1 & x > 0
}

x_ok(NULL)
x_ok(1)
x_ok(1:3)

```

Runs each logical test, no control flow. Provide values to tests that should not accept them

2. What does this function return? Why? Which principle does it illustrate?

```{r}
f2 <- function(x = z) {
  z <- 100
  x
}
f2()
```

100, lazy evaluation

3. What does this function return? Why? Which principle does it illustrate?

```{r}
y <- 10
f1 <- function(x = {y <- 1; 2}, y = 0) {
  c(x, y)
}
f1()
y
```

(2, 1), 10. name masking

4. In hist(), the default value of xlim is range(breaks), the default value for breaks is "Sturges", and

```{r}
range("Sturges")
```
Explain how hist() works to get a correct xlim value.

Sturges is a function called which calculates bins for you

5. Explain why this function works. Why is it confusing?

```{r}
show_time <- function(x = stop("Error!")) {
  stop <- function(...) Sys.time()
  print(x)
}
show_time()
```

The default value for x is run the function stop with the argument "Error!". Which should stop and prevent an error message. The stop function is then redefined in the function to provide the system time instead. They evaluation is then applied to the value of x. Last x is print given its environmental value which is the system time. It's confusing because we are masking functions and lazy evaluating calls.

6. How many arguments are required when calling library()?

None