---
title: "9_Functionals"
output: html_document
date: "2023-01-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 9. Functionals

## 9.1 Introduction

```{r}
randomise <- function(f) f(runif(1e3))
randomise(mean)
randomise(mean)
randomise(sum)
```

```{r}
library(purrr)
```

## 9.2 My first functional: `map()`

```{r}
triple <- function(x) x * 3
map(1:3, triple)
```

```{r}
simple_map <- function(x, f, ...) {
  out <- vector("list", length(x))
  for (i in seq_along(x)) {
    out[[i]] <- f(x[[i]], ...)
  }
  out
}
```

### 9.2.1 Producing atomic vectors

```{r}
map_chr(mtcars, typeof)
```

```{r}
map_lgl(mtcars, is.double)
```

```{r}
n_unique <- function(x) length(unique(x))
map_int(mtcars, n_unique)
```

```{r}
map_dbl(mtcars, mean)
```

```{r, error = T}
pair <- function(x) c(x, x)
map_dbl(1:2, pair)
```

```{r, error = T}
map_dbl(1:2, as.character)
```

```{r}
map(1:2, pair)
```

```{r}
map(1:2, as.character)
```

```{r, error = T}
map_dbl(x, mean, na.rm = TRUE)

=====
  
vapply(x, mean, na.rm = TRUE, FUN.VALUE = double(1))
```

### 9.2.2 Anonymous functions and shortcuts

```{r}
map_dbl(mtcars, function(x) length(unique(x)))
```

```{r}
map_dbl(mtcars, ~ length(unique(.x)))
```

```{r}
as_mapper(~ length(unique(.x)))
```

```{r}
x <- map(1:3, ~ runif(2))
str(x)
```

```{r}
x <- list(
  list(-1, x = 1, y = c(2), z = "a"),
  list(-2, x = 4, y = c(5, 6), z = "b"),
  list(-3, x = 8, y = c(9, 10, 11))
)
```

```{r}
map_dbl(x, "x")
```

```{r}
map_dbl(x, 1)
```

```{r}
map_dbl(x, list("y", 1))
```

```{r, error = T}
map_chr(x, "z")
```

```{r}
map_chr(x, "z", .default = NA)
```

### 9.2.3 Passing arguments with `...`

```{r}
x <- list(1:5, c(1:10, NA))
map_dbl(x, ~ mean(.x, na.rm = TRUE))
```

```{r}
map_dbl(x, mean, na.rm = TRUE)
```

```{r}
plus <- function(x, y) x + y

x <- c(0, 0, 0, 0)
map_dbl(x, plus, runif(1))
```

```{r}
map_dbl(x, ~ plus(.x, runif(1)))
```

### 9.2.4 Argument names

```{r, error = T}
boostrap_summary <- function(x, f) {
  f(sample(x, replace = TRUE))
}

simple_map(mtcars, boostrap_summary, f = mean)
```

### 9.2.5 Varying another argument

```{r}
trims <- c(0, 0.1, 0.2, 0.5)
x <- rcauchy(1000)
```

```{r}
map_dbl(trims, ~ mean(x, trim = .x))
```

```{r}
map_dbl(trims, function(trim) mean(x, trim = trim))
```

```{r}
map_dbl(trims, mean, x = x)
```

#### 9.2.6 Exercises

1.  Use `as_mapper()` to explore how purrr generates anonymous functions for the integer, character, and list helpers. What helper allows you to extract attributes? Read the documentation to find out.

```{r}
as_mapper(c("a", "b", "c"))
```

```{r}
as_mapper(c(1, 2, 3))
```

```{r}
as_mapper(list(1, "a", 2))
```

```{r}
as_mapper(list(1, attr_getter("a")))
```

Looks like it is using `pluck_raw`. Get attributes wiht `attr_getter`

2.  `map(1:3, ~ runif(2))` is a useful pattern for generating random numbers, but `map(1:3, runif(2))` is not. Why not? Can you explain why it returns the result that it does?

```{r}
map(1:3, ~ runif(2))
```

```{r}
map(1:3, runif(2))
```

First one creates an anonymous function which then generates 2 random numbers for each of the 3 iterations. The second only generates one set of random numbers which is fed into map and results in maps default values being spit out when piped to `as_mapper`

3.  Use the appropriate `map()` function to:

-   Compute the standard deviation of every column in a numeric data frame

```{r}
mat <- as.data.frame(matrix(1:25, nrow = 5))
map_dbl(mat, ~ sd(.x))
```

-   Compute the standard deviation of every numeric column in a mixed data frame. (Hint: you'll need to do it in two steps.)

```{r}
summary(iris)
map_dbl(iris[map_lgl(iris, is.numeric)], ~ sd(.x))
```

-   Compute the number of levels for every factor in a data frame

```{r}
library(tidyverse)
summary(attenu)
df <- attenu %>%
  mutate(event = as.factor(event))
summary(df)

map_int(df[map_lgl(df, ~ is.factor(.x))], ~ length(levels(.x)))

```

4.  The following code simulates the performance of a t-test for non-normal data. Extract the p-value from each test, then visualise.

```{r}
trials <- map(1:100, ~ t.test(rpois(10, 10), rpois(7, 10)))

map_dbl(trials, "p.value") %>%
  hist(main = "Histogram of p values")
```

5.  The following code uses a map nested inside another map to apply a function to every element of a nested list. Why does it fail, and what do you need to do to make it work?

```{r, error = T}
x <- list(
  list(1, c(3, 9)),
  list(c(3, 6), 7, c(4, 7, 6))
)

triple <- function(x) x * 3
map(x, map, .f = triple)

map(x, map, triple)
# or
map(x, ~ map(.x, triple))
```

Using .f makes triple the function of the outer map call and not the inner map call. Just remove the name and it will go in order and work

6.  Use `map()` to fit linear models to the mtcars dataset using the formulas stored in this list:

```{r}
formulas <- list(
  mpg ~ disp,
  mpg ~ I(1 / disp),
  mpg ~ disp + wt,
  mpg ~ I(1 / disp) + wt
)

map(formulas, lm, data = mtcars)
```

7.  Fit the model `mpg ~ disp` to each of the bootstrap replicates of `mtcars` in the list below, then extract the $R^2$ of the model fit (Hint: you can compute the $R^2$ with summary().)

```{r}
bootstrap <- function(df) {
  df[sample(nrow(df), replace = TRUE), , drop = FALSE]
}

bootstraps <- map(1:10, ~ bootstrap(mtcars))
head(bootstraps)

map(bootstraps, ~ lm(mpg ~ disp, data = .x)) %>%
  map(summary) %>%
  map("r.squared")
```

## 9.3 Purrr style